Factors affecting health sector expenditure in OECD countries

### 2025-04-27 products/manuscript/Manuscript.qmd
intro #1

>>>>

Health expenditure is an important indicator of a country's committment to citizen's well-being and efficiency of it's healthcare systems. Among the OECD countries, the per capita health expenditure vary significantly ranging from 1,000 USD to over 10,000 [@2021].
Understanding the factors that drive these differences is important for policymakers to optimize healthcare delivery and improving healhcare access. While GDP had been one of the primary determinant of health expending [@Newhouse1977], recent studies suggest that institutional, demographic, and socio-economic factors may also play a crucial role in determination of health expenditure among different countries [@Ziegler2017].

The relationship between health expenditure and their determinants because particularly relevant in the context of recent global challenges. The recent COVID-19 pandemic also exposed the vulnerabilities in global healthcare systems with countries showing high pre-pandemic health expenditure generally demonstrating greater resilience [@WHO2022]
